---
title: "Trend Estimation - 1"
author: "Johnathan McCabe"
date: "Monday, April 20, 2015"
output: html_document
---



# Data Exploration

Know that all losses are fully developed so investigate quantiles and mean.

Simple average

```{r}
library(dplyr)
library(ggplot2)
library(reshape2)

set.1.df <- read.csv("data_set_1.csv")

probs <- c(0.1,0.25,0.5,0.75,0.9)

years <- 1:max(set.1.df$year)

quantile.results <- do.call("rbind", lapply(years, function(yr){
  
  df <- set.1.df %>% filter(year == yr)
  q<-  quantile(df$loss, probs = probs)
  
  q
}))

quantile.results <- as.data.frame(quantile.results, stringsAsFactors=F)
quantile.results$year <- as.numeric(row.names(quantile.results))

res <- quantile.results %>%
  melt(value.name = "loss", variable.name = "type",id.vars = "year")

means <- set.1.df %>%
  group_by(year) %>%
  summarise(loss = mean(loss)) %>%
  mutate(type = "mean")


res <- res %>%
  rbind(means) %>%
  mutate(year_factor = factor(year))


p <- ggplot(res, aes(x = year_factor,y =  loss))
p <- p + geom_line(aes(color = type, group = type), stat="identity") +
  theme(legend.title = element_blank()) +
  xlab("Year") +
  ylab("Loss")

print(p)
```

There is a definite general upward trend.

# Linear Model Fit to Log Mean Loss

Try fitting a linear model to the log of the means.

```{r}

log_means <- means %>% 
  mutate(log_mean = log(loss))

mod <- lm(data = log_means, formula = log_mean ~ year)

mod$coefficients

initial_mean <- exp(mod$coefficients[1])
trend <- exp(mod$coefficients[2]) - 1


```

So the the initial mean for year 1 is `r round(initial_mean,0)` with trend rate `r sprintf("%.2f%%", trend*100)`

```{r}

x<-1:10
y<-initial_mean * (1+trend)^(x )
p2 <- p + geom_line(data = data.frame(x,y), aes(x,y), size=1)

print(p2)

summary(mod)
```


R^2^ is quite large so seems a decent fit.


# Linear Model Fit to Log Loss

Try fitting a linear model to the log of the losses.

```{r}

log_losses <- set.1.df %>% 
  mutate(log_loss = log(loss))

mod.losses <- lm(data = log_losses, formula = log_loss ~ year)

mod.losses$coefficients

initial_mean <- exp(mod.losses$coefficients[1])
trend <- exp(mod.losses$coefficients[2]) - 1


```

So the the initial mean for year 1 is `r round(initial_mean,0)` with trend rate `r sprintf("%.2f%%", trend*100)`

```{r}

x<-1:10
y<-initial_mean * (1+trend)^(x )
p3 <- p + geom_line(data = data.frame(x,y), aes(x,y), size=1)

print(p3)

summary(mod.losses)
```


R^2^ is quite large so again seems a decent fit.

Results are very similar (`r sprintf("%.2f%%", trend *100)` vs `r sprintf("%.2f%%", (exp(mod$coefficients[2]) - 1) *100)`)

