---
title: "Trend Estimation - 1"
author: "Johnathan McCabe"
date: "Monday, April 20, 2015"
output: html_document
---
  
  
  
# Data Exploration
  
  Know that all losses are fully developed so investigate quantiles and mean.

Simple average

```{r data-explore}
library(dplyr)
library(ggplot2)
library(reshape2)
library(runjags)
library(coda)
library(knitr)

data.2 <- read.csv("../data_set_2.csv", stringsAsFactors=F)



quantile_results <- function(data){
  years <- unique(data.2$year)
  
  retVal = do.call("rbind", lapply(years,function(yr){
    
    year_losses <- data %>% filter(year == yr) %>% select(loss) %>% unlist()
    
    quantiles <- quantile(year_losses,probs = c(0.1,0.25,0.5,0.75,0.9))
    
    result <- data.frame(as.list(quantiles))
    result$year <- yr
    
    return(result)
  }))
  
  return(retVal)
}


x <- quantile_results(data.2)


means <- data.2 %>%
  group_by(year) %>%
  summarise(loss = mean(loss)) 


plot.data <- x %>% melt(id.vars = c("year"), variable.name = "quantile", value.name = "loss") %>%
  rbind(means %>% mutate(quantile = "mean"))

p <- ggplot(plot.data, aes(year, loss))
p <- p + geom_line(aes(color=quantile))
p

lmeans <- means %>% mutate(lmean = log(loss)) %>% select(-loss)

m <- lm(data = lmeans, formula = lmean ~ year)


print(p)
```

There is a definite general upward trend.

# Linear Model Fit to Log Mean Loss

Try fitting a linear model to the log of the means.

```{r lm-log-mean}

log_means <- means %>% 
  mutate(log_mean = log(loss))

mod <- lm(data = log_means, formula = log_mean ~ year)

mod$coefficients

initial_mean <- exp(mod$coefficients[1])
trend <- exp(mod$coefficients[2]) - 1


```

So the the initial mean for year 1 is `r round(initial_mean,0)` with trend rate `r sprintf("%.2f%%", trend*100)`

```{r}

x<-1:10
y<-initial_mean * (1+trend)^(x )
p2 <- p + geom_line(data = data.frame(x,y), aes(x,y), size=1)

print(p2)

summary(mod)
```


R^2^ is quite large so seems a decent fit.


# Linear Model Fit to Log Loss

Try fitting a linear model to the log of the losses.

```{r lm}

log_losses <- data.2 %>% 
  mutate(log_loss = log(loss))

mod.losses <- lm(data = log_losses, formula = log_loss ~ year)

mod.losses$coefficients

initial_mean <- exp(mod.losses$coefficients[1])
trend <- exp(mod.losses$coefficients[2]) - 1


```

So the the initial mean for year 1 is `r round(initial_mean,0)` with trend rate `r sprintf("%.2f%%", trend*100)`

```{r}

x<-1:10
y<-initial_mean * (1+trend)^(x )
p3 <- p + geom_line(data = data.frame(x,y), aes(x,y), size=1)

print(p3)

summary(mod.losses)
```


R^2^ is quite large so again seems a decent fit.

Results are very similar (`r sprintf("%.2f%%", trend *100)` vs `r sprintf("%.2f%%", (exp(mod$coefficients[2]) - 1) *100)`)


# Empirical Bayes

Three bases


```{r bayes.1}
####################################################################################
# JAGS script 
####################################################################################
modelString = "model {
for (i in 1:370){
tmu[i] <- mu + ltrend*(year[i] - 1)    #mu is the log mean for year 8 

lclaims[i] ~ dnorm(tmu[i],1/sigma^2)

}
#
# specify priors
#
ltrend ~ dunif(-1,1)
mu ~ dunif(0,35)
sigma ~ dunif(0,1)
}" 
#
# initialize JAGS model
#
inits1=list(.RNG.name= "base::Wichmann-Hill",
            .RNG.seed= 12341)
inits2=list(.RNG.name= "base::Marsaglia-Multicarry",
            .RNG.seed= 12342)
inits3=list(.RNG.name= "base::Super-Duper",
            .RNG.seed= 12343)
inits4=list(.RNG.name= "base::Mersenne-Twister",
            .RNG.seed= 12344)

data.for.jags=list(lclaims = log(data.2$loss),
                   year = data.2$year,
                   loss = data.2$loss)
#
# run the model
# #

jagout=run.jags(model=modelString,monitor=c("ltrend","mu","sigma"),
                data=data.for.jags,n.chains=4,method="parallel",
                inits=list(inits1,inits2,inits3,inits4),thin=4,
                plots=TRUE,burnin=5000,sample=2500,psrf.target=1.05)
print(jagout$timetaken)

plot(jagout)

gelman=gelman.diag(jagout)
maxpsrf=max(gelman$psrf[,1])
print(paste("maxpsrf =",maxpsrf))

samples <- as.matrix(jagout$mcmc)

jag.means <- sapply(1:10, function(i){
  exp(mean(samples[,2] + samples[,1] * (i - 1)))
})
df.lnorm <- data.frame(x=1:10,y=jag.means,quantile="lnorm.jag.mean")
jag.means

p_lnorm <- p + geom_line(data = df.lnorm, aes(x,y))
print(p_lnorm)

jagout$summary


########################################################################################
# GAMMA MODEL
########################################################################################

mean <- mean(data.2$loss)
sd <- sd(data.2$loss)

shape <- sd*sd / mean
rate <- shape/mean
####################################################################################
# JAGS script 
####################################################################################
modelString.gamma = "model {
for (i in 1:370){

rt[i] <- rate/(trend + 1)^(year[i] - 1)
loss[i] ~ dgamma(shape, rt[i])

}

trend ~ dunif(-1,1)
shape ~ dunif(100,10000)
rate ~ dunif(0.000001,0.25)
}" 
#
# initialize JAGS model
#
inits1=list(.RNG.name= "base::Wichmann-Hill",
            .RNG.seed= 12341)
inits2=list(.RNG.name= "base::Marsaglia-Multicarry",
            .RNG.seed= 12342)
inits3=list(.RNG.name= "base::Super-Duper",
            .RNG.seed= 12343)
inits4=list(.RNG.name= "base::Mersenne-Twister",
            .RNG.seed= 12344)

data.for.jags=list(lclaims = log(data.2$loss),
                   year = data.2$year,
                   loss = data.2$loss)
#
# run the model
# #

jagout.gamma =run.jags(model=modelString.gamma,monitor=c("rate","shape","trend"),
                       data=data.for.jags,n.chains=4,method="parallel",
                       inits=list(inits1,inits2,inits3,inits4),thin=4,
                       plots=TRUE,burnin=5000,sample=2500,psrf.target=1.05)
print(jagout.gamma$timetaken)


plot(jagout.gamma)

gelman=gelman.diag(jagout.gamma)
maxpsrf=max(gelman$psrf[,1])
print(paste("maxpsrf =",maxpsrf))

samples <- as.matrix(jagout.gamma$mcmc)


jag.means <- sapply(1:10, function(i){
  mean(samples[,2] * (1+samples[,3]) ^ (i - 1) / samples[,1])
})
df.gamma <- data.frame(x=1:10,y=jag.means,quantile="gamma.jag.mean")

p2 <- p_lnorm +geom_line(data = df.gamma, aes(x,y))
print(p2)


########################################################################################
# UNIFORM MODEL
########################################################################################

####################################################################################
# JAGS script 
####################################################################################
modelString.unif = "model {
for (i in 1:370){

tmu[i] <- mu + ltrend * (year[i] - 1)
lclaims[i] ~ dunif(tmu[i] - 4 * sig, tmu[i] + 4 * sig)

}

ltrend ~ dunif(-1,1)
sig ~ dlnorm(0.0001,2)
mu ~ dunif(7,15)
}" 
#
# initialize JAGS model
#
inits1=list(.RNG.name= "base::Wichmann-Hill",
            .RNG.seed= 12341)
inits2=list(.RNG.name= "base::Marsaglia-Multicarry",
            .RNG.seed= 12342)
inits3=list(.RNG.name= "base::Super-Duper",
            .RNG.seed= 12343)
inits4=list(.RNG.name= "base::Mersenne-Twister",
            .RNG.seed= 12344)

data.for.jags=list(lclaims = log(data.2$loss),
                   year = data.2$year,
                   loss = data.2$loss)
#
# run the model
# #

jagout.unif =run.jags(model=modelString.unif,monitor=c("ltrend","mu","sig"),
                      data=data.for.jags,n.chains=4,method="parallel",
                      inits=list(inits1,inits2,inits3,inits4),thin=4,
                      plots=TRUE,burnin=10000,sample=2500,psrf.target=1.05)
print(jagout.unif$timetaken)


plot(jagout.unif)

gelman=gelman.diag(jagout.unif)
maxpsrf=max(gelman$psrf[,1])
print(paste("maxpsrf =",maxpsrf))

samples <- as.matrix(jagout.unif$mcmc)


jag.unif.means <- sapply(1:10, function(i){
  exp(mean(samples[,2] + samples[,1] * (i - 1)))
})
df.unif <- data.frame(x=1:10,y=jag.unif.means,quantile="unif.jag.mean")

p3 <- p2 + geom_line(data = df.unif, aes(x,y))
print(p3)
```

```{r results='asis'}

means <- c(exp(jagout$summary$statistics[1,1]) - 1,
           jagout.gamma$summary$statistics[3,1],
           exp(jagout.unif$summary$statistics[1,1])-1)

df <- data.frame(method = c("Lognormal", "Gamma", "Uniform"), value = means)
k <- kable(df)

print(k)


```


# Empirical Bayes - Trend Estimate By Year
```{r bayes.2}

####################################################################################
# JAGS script 
####################################################################################
modelString = "model {
for (i in 1:length(lclaims)){

 log_trend[i,1] <- 0
for(j in 2:year[i]){
  log_trend[i,j] <- log_trend[i,j-1] + ltrend[j]
}
tmu[i] <- mu + log_trend[i,year[i]]   #mu is the log mean for year 8 

lclaims[i] ~ dnorm(tmu[i],1/sigma^2)

}
#
# specify priors
#
ltrend[1] <- 0
for(i in 2:num_years){
  ltrend[i] ~ dnorm(-1,1)
}
mu ~ dunif(0,35)
sigma ~ dunif(0,1)
}" 
#
# initialize JAGS model
#
inits1=list(.RNG.name= "base::Wichmann-Hill",
            .RNG.seed= 12341)
inits2=list(.RNG.name= "base::Marsaglia-Multicarry",
            .RNG.seed= 12342)
inits3=list(.RNG.name= "base::Super-Duper",
            .RNG.seed= 12343)
inits4=list(.RNG.name= "base::Mersenne-Twister",
            .RNG.seed= 12344)

data.for.jags=list(lclaims = log(data.2$loss),
                   year = data.2$year,
                   loss = data.2$loss,
                   num_years = max(data.2$year))
#
# run the model
# #

jagout=run.jags(model=modelString,monitor=c("ltrend","mu","sigma"),
                data=data.for.jags,n.chains=4,method="parallel",
                inits=list(inits1,inits2,inits3,inits4),thin=10,
                plots=TRUE,burnin=10000,sample=10000,psrf.target=1.05)
print(jagout$timetaken)


samples <- as.data.frame(as.matrix(jagout$mcmc))

samples$cum_1 <- 0 
for(j in 2:10){
  col <- paste("cum_",j,sep="")
  prev_col <- paste("cum_",j-1,sep="")
  samples[,col] <- samples[,prev_col] + samples[,j]
}


jag.means <- sapply(1:10, function(i){

  exp(mean(samples[,11] + samples[,12 + i]))
})

df.lnorm <- data.frame(x=1:10,y=jag.means,quantile="lnorm.jag.mean")
jag.means

p_lnorm.2 <- p + geom_line(data = df.lnorm, aes(x,y))
print(p_lnorm.2)

```


```{r results='asis'}

plot(jagout)
```
